### 2020-12-03 arxiv/tools.tex

\begin{corollary} \label{cor_largedeviation}
For any deterministic vector $v\in \R^p$, we have that w.h.p.
\be\label{Zv_cor}\left|\|Zv\|^2- n\|v\|^2\right|\le  n^{1-c_\varphi} \|v\|^2.\ee
\end{corollary}
\begin{proof}
	Let $Q= n^{ \frac{2}{\varphi}}\log n$.
	We introduce a truncated matrix $\wt Z$ with entries $\wt Z_{ij}:= \mathbf 1\left( |Z_{ij}|\le Q \right)\cdot Z_{ij}$. Then $\wt Z$ is equal to $Z$ when all the entries of $Z$ are smaller than $Q$. Using Markov's inequality and a simple union bound (cf. equation \eqref{Ptrunc}), this happens with probability 
\begin{equation}\label{XneX222}
\mathbb P(\wt Z= Z) =1-\OO ( (\log n)^{-\varphi}).
\end{equation}
Furthermore, using the finite $\varphi$-th moment condition and the tail probabilities, we can show that the mean and variance of $Z-\wt Z$ are small, which gives that (cf. equation \eqref{meanshif}) %, we have that
\be\label{meanshif222}
	|\mathbb E  \wt  Z_{ij}| =\OO(n^{-3/2}), \quad  \mathbb E |\wt  Z_{ij}|^2 =1+ \OO(n^{-1}).
\ee
%From the first estimate in equation \eqref{meanshif}, we can also get a bound on the operator norm:
%\be\label{EZ norm222}\|\E \wt Z^{(1)}\|=\OO(n^{-1/2}) .
%\ee
We centralize and rescale $\wt Z$ as $ \wh Z:=\frac{\wt Z - \E \wt Z }{ (\E|\wt Z_{11}|^2)^{1/2}}.$
It suffices to show that 
\be\label{show_whZ}
 \left|\|\wh Zv\|^2- n\|v\|^2\right| \le n^{1/2 + \e} Q\|v\|^2.
\ee 
This is because provided with equation \eqref{show_whZ}, using \eqref{meanshif222} we can get that
$$ \left|\|\wt Zv\|^2- n\|v\|^2\right| \le n^{1/2 + \e} Q\|v\|^2,$$
which implies \eqref{Zv_cor} for any fixed value $c_{\varphi}$ within $(0, \frac{\varphi - 4}{2\varphi})$.

To prove equation \eqref{show_whZ}, we first observe that for $1\le i \le n$, $(\wh Zv)_i=\sum_{1\le j \le p} \wh Z_{ij}v_j$ are i.i.d random variables of mean zero and variance $\|v\|^2$. Moreover, using equation \eqref{eq largedev10} we get that with overwhelming probability
$$ |(\wh Zv)_i| \le n^\e Q\max_{1\le i\le p}|v_i| + n^\e \|v\| . $$
Together with the trivial bound $\max_{1\le i\le p}|v_i| \le \|v\|$, this implies that the random variables $\frac{(\wh Zv)_i}{\|v\|}$, $1\le i \le p$, are i.i.d random variables of mean zero, variance 1 and bounded support $Q$. Then applying equation \eqref{eq largedev20}, we get that
$$ \left|\|\wh Zv\|^2 - n\|v\|^2\right|=\left|\sum_i \left(|(\wh Zv)_i|^2-\E|(\wh Zv)_i|^2\right)\right|\le n^{1/2+\e} Q  \|v\|^2 .
 $$
This concludes equation \eqref{show_whZ}.
 \end{proof}

@@

\begin{corollary}\label{cor_calE}
Let $\e>0$ be a sufficiently small fixed constant.
For any deterministic vector $v\in \R^n$, we have that w.h.p.
\be\label{vcalE}
	|v^\top \epsilon^{(i)}| \le n^{c} \cdot \sigma \|v\|, \text{ and } \|v^\top \cal E\| \le n^{c} \cdot \sigma  \|v\|.
\ee
For any deterministic matrix $B \in \real^{n\times n}$, we have that w.h.p.
\begin{align}\label{vcalA2}
	\left|(\epsilon^{(i)})^\top B \epsilon^{(j)} - \delta_{ij}\cdot \sigma^2 \tr(B)\right| \le n^{c}\cdot \sigma^2 \|B\|_F, \text{ and }
	\left\| \cal E^\top B \cal E - \sigma^2 \tr[B]\cdot \id_{t\times t}\right\|_F \le n^{c} \cdot  \sigma^2\|B\|_F.
\end{align}
%\be\label{vcalAE3}
%,
%\ee
\end{corollary}
\begin{proof}
Rescaling $\epsilon^{(i)}$ by $\sigma$, we get a random vector with zero mean, unit variance, and bounded moments up to any order.
Using the first estimate in equation \eqref{eq largedev1} of Lemma \ref{largedeviation} (recall that the stochastic domination notation means the inequality holds with a multiplicative factor of $p^{\e}$ on the right), we obtain that equation \eqref{vcalE} holds (the second result is a consequence of the first).
Using the second estimate in equation \eqref{eq largedev1}, we obtain that for $i\ne j$,
 \begin{align*}
\left|(\epsilon^{(i)})^\top B \epsilon^{(j)}\right| = \Big|\sum_{k,l=1}^n \epsilon^{(i)}_k\epsilon^{(j)}_l B_{kl} \Big| \le p^{\e} \cdot \sigma^2 \Big(\sum_{k,l=1}^n|B_{kl}|^2\Big)^{1/2}=\sigma^2 \|B\|_F.
 \end{align*}
Using the two estimates in equation \eqref{eq largedev2}, we obtain that 
\begin{align*}
\left|(\epsilon^{(i)})^\top B \epsilon^{(i)} - \sigma^2 \tr[B]\right| &\le \Big| \sum_{k=1} \left(|\epsilon^{(i)}_k|^2-\E|\epsilon^{(i)}_k|^2\right)B_{ii} \Big|+\Big| \sum_{ k\ne l } \epsilon^{(i)}_k \epsilon^{(i)}_l B_{kl} \Big| \\
&\le p^{\e}\cdot \sigma^2 \Big( \sum_{k}|B_{kk}|^2\Big)+ p^{\e}\cdot \sigma^2 \Big( \sum_{k\ne l}|B_{kl}|^2\Big)^{1/2}\le n^{\e}\cdot\sigma^2 \|B\|_F.
\end{align*}
Hence, we have shown equation \eqref{vcalA2} (the second result is again a consequence of the first result).
\end{proof}

>>>>

\begin{corollary} \label{cor_largedeviation}
For any deterministic vector $v\in \R^p$, we have that w.h.p.
\be\label{Zv_cor}\left|\|Zv\|^2- n\|v\|^2\right|\le 2n^{1-c_\varphi} \|v\|^2.\ee
\end{corollary}
\begin{proof}
	Let $Q= n^{ \frac{2}{\varphi}}\log n$.
	We introduce a truncated matrix $\wt Z$ with entries $\wt Z_{ij}:= \mathbf 1\left( |Z_{ij}|\le Q \right)\cdot Z_{ij}$. Then $\wt Z$ is equal to $Z$ when all the entries of $Z$ are smaller than $Q$. Using Markov's inequality and a simple union bound (cf. equation \eqref{Ptrunc}), this happens with probability 
\begin{equation}\label{XneX222}
\mathbb P(\wt Z= Z) =1-\OO ( (\log n)^{-\varphi}).
\end{equation}
Furthermore, using the finite $\varphi$-th moment condition and the tail probabilities, we can show that the mean and variance of $Z-\wt Z$ are small, which gives that (cf. equation \eqref{meanshif}) %
\be\label{meanshif222}
	|\mathbb E  \wt  Z_{ij}| =\OO(n^{-3/2}), \quad  \mathbb E |\wt  Z_{ij}|^2 =1+ \OO(n^{-1}).
\ee
We centralize and rescale $\wt Z$ as $ \wh Z:=\frac{\wt Z - \E \wt Z }{ (\E|\wt Z_{11}|^2)^{1/2}}.$
Let $\e$ be a sufficiently small fixed constant.
To prove the result, it suffices to show that
\be\label{show_whZ}
 \left|\|\wh Zv\|^2- n\|v\|^2\right| \le n^{1/2 + \e} Q\|v\|^2.
\ee 
This is because provided with equation \eqref{meanshif222} and \eqref{show_whZ}, we can get that
$$ \left|\|\wt Zv\|^2- n\|v\|^2\right| \le n^{1/2 + \e} Q\|v\|^2,$$
which implies the desired our (recall that $c_{\varphi} < 1/2 - \e$).
To prove equation \eqref{show_whZ}, we first show that for any $i = 1, 2, \dots, n$, $(\wh Zv)_i=\sum_{1\le j \le p} \wh Z_{ij}v_j$ is at most $2n^{\e} Q$.
Since $\wh{Z} v$ consists of i.i.d random variables with mean zero and variance $\|v\|^2$,
using equation \eqref{eq largedev10} from Lemma \ref{largedeviation}, we get that with overwhelming probability
$$ |(\wh Zv)_i| \le n^\e Q\max_{1\le i\le p}|v_i| + n^\e \|v\| \le 2n^{\e} Q. $$
Hence, $\frac{(\wh Zv)}{\|v\|}$ consists of i.i.d random variables with mean zero, unit variance, and bounded support $2n^{\e}Q$.
Applying equation \eqref{eq largedev20}, we get that
$$ \left|\|\wh Zv\|^2 - n\|v\|^2\right|=\left|\sum_i \left(|(\wh Zv)_i|^2-\E|(\wh Zv)_i|^2\right)\right|\le 2n^{1/2+2\e} Q  \|v\|^2 .
 $$
Hence the proof is complete.
 \end{proof}

@@

\begin{corollary}\label{cor_calE}
Let $\e>0$ be a sufficiently small fixed constant.
For any deterministic vector $v\in \R^n$, we have that w.h.p.
\be\label{vcalE}
	|v^\top \epsilon^{(i)}| \le n^{c} \cdot \sigma \|v\|, \ \ \text{ and }\ \ \|v^\top \cal E\| \le n^{c} \cdot \sigma  \|v\|.
\ee
For any deterministic matrix $B \in \real^{n\times n}$, we have that w.h.p.
\begin{align}\label{vcalA2}
	\left|(\epsilon^{(i)})^\top B \epsilon^{(j)} - \delta_{ij}\cdot \sigma^2 \tr(B)\right| \le n^{c}\cdot \sigma^2 \|B\|_F,  \ \ \text{ and }\ \
	\left\| \cal E^\top B \cal E - \sigma^2 \tr[B]\cdot \id_{t\times t}\right\|_F \le n^{c} \cdot  \sigma^2\|B\|_F.
\end{align}
\end{corollary}
\begin{proof}
Rescaling $\epsilon^{(i)}$ by $\sigma$, we get a random vector with zero mean, unit variance, and bounded moments up to any order.
Using the first estimate in equation \eqref{eq largedev1} of Lemma \ref{largedeviation} (recall that the stochastic domination notation means the inequality holds with a multiplicative factor of $p^{\e}$ on the right), we obtain that equation \eqref{vcalE} holds (the second result is a consequence of the first).
Using the second estimate in equation \eqref{eq largedev1}, we obtain that for $i\ne j$,
 \begin{align*}
\left|(\epsilon^{(i)})^\top B \epsilon^{(j)}\right| = \Big|\sum_{k,l=1}^n \epsilon^{(i)}_k\epsilon^{(j)}_l B_{kl} \Big| \le p^{\e} \cdot \sigma^2 \Big(\sum_{k,l=1}^n|B_{kl}|^2\Big)^{1/2}=\sigma^2 \|B\|_F.
 \end{align*}
Using the two estimates in equation \eqref{eq largedev2}, we obtain that 
\begin{align*}
\left|(\epsilon^{(i)})^\top B \epsilon^{(i)} - \sigma^2 \tr[B]\right| &\le \Big| \sum_{k=1} \left(|\epsilon^{(i)}_k|^2-\E|\epsilon^{(i)}_k|^2\right)B_{ii} \Big|+\Big| \sum_{ k\ne l } \epsilon^{(i)}_k \epsilon^{(i)}_l B_{kl} \Big| \\
&\le p^{\e}\cdot \sigma^2 \Big( \sum_{k}|B_{kk}|^2\Big)+ p^{\e}\cdot \sigma^2 \Big( \sum_{k\ne l}|B_{kl}|^2\Big)^{1/2}\le n^{\e}\cdot\sigma^2 \|B\|_F.
\end{align*}
Hence, we have shown equation \eqref{vcalA2} (the second result is again a consequence of the first result).
\end{proof}

### 2020-12-03 old/arxiv_full_version/experiments_text.tex

\paragraph{Sample size ratio.}
First, we show that our observation in Figure \ref{fig_size} also occurs in the text classification tasks.
In Figure \ref{fig_ab_data}, we observe that for multiple example task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually.
On the $y$-axis, we plot task two's test accuracy using HPS, subtracted by its STL test accuracy.
%validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
We fix task two's sample size at $1000$ and increase task one's sample size from $100$ to $3000$.

These examples and the one in Figure \ref{fig_size} suggest a natural progressive training schedule, where we add samples progressively until performance drops.
Concretely, here is one implementation of this idea.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
%Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the training data into $S$ batches.
	We divide the training procedure into $S$ stages. During every stage, we progressively add one more data batch.
	\item During every stage, we train for $T$ epochs using only the $S$ batches. If the validation accuracy drops compared to the previous round's result or reach a desired threshold $\tau$, we terminate.
%	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}
For example, if we apply this procedure to the settings of Figure \ref{fig_ab_data} and \ref{fig_size}, it will terminate once reaching the optimal sample ratio.
%See Algorithm \ref{alg_inc_train} for a complete description for two tasks.
The advantage of this procedure is that it reduces the computational cost compared to standard round-robin training schedules.
For example, if the procedure terminates at 30\% of all batches, then SGD only passes over 30\% of its data, whereas standard round-robin training passes over 100\% of task one's data.

>>>>

\paragraph{Sample size ratio.}
First, we show that our observation in Figure \ref{fig_size} also occurs in the text classification tasks.
In Figure \ref{fig_ab_data}, we observe that for multiple example task pairs, increasing task one's sample size improves task two's prediction accuracy initially but hurts eventually.
On the $y$-axis, we plot task two's test accuracy using HPS, subtracted by its STL test accuracy.
%validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
We fix task two's sample size at $1000$ and increase task one's sample size from $100$ to $3000$.

These examples and Figure \ref{fig_size} suggest a natural progressive training schedule, where we add samples progressively until performance drops.
Concretely, here is one implementation of this idea.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
%Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the training data into $S$ batches.
	We divide the training procedure into $S$ stages. During every stage, we progressively add one more data batch.
	\item During every stage, we train for $T$ epochs using only the $S$ batches. If the validation accuracy drops compared to the previous round's result or reaches the desired threshold $\tau$, we terminate.
%	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}
If we apply this procedure to the settings of Figure \ref{fig_ab_data} and \ref{fig_size}, it will terminate once reaching the optimal sample ratio.
%See Algorithm \ref{alg_inc_train} for a complete description for two tasks.
The advantage of this procedure is that it reduces the computational cost compared to standard round-robin training schedules.
For example, if the process terminates at 30\% of all batches, then SGD only passes over 30\% of its data, whereas standard round-robin training passes over 100\% of task one's data.
